Disc stuff

### 2012-10-31 paper.tex
spell check

%-describe the envelope structure (pre-stellar core)
The model used to simulate the emission from a young protoplantary disc is a hybrid model created from hydrodynamical models of an infalling prestellar core and a gravitationally unstable protoplanetary disc. The prestellar core model is based upon the model of Keto \& Caselli (2010). This describes an infalling pre-stellar core, with densities temperatures and velocities given by the model of the collapse of a 10 solar mass Bonnor-Ebert sphere and provides a match to line profiles seen in the pre-stellar core L1544. Some slight modification due to the addition of Oxygen lines (Keto et al 2012 in prep) have been been added to this model resulting in a slight decrease in the peak infall velocity. This prestellar core model extends from 80$\,$au out to 10,000$\,$AU. The model is a 1D spherically symmetric model with inward motions (see figure \ref{l1544model}).\newline

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/rhoT2.png}
 \label{rhoT} 
 \caption{Left: A 3D plot of log density (m$^3$) showing the spiral structure in the xy plane and scale height of the disc. Right: The 3D temperature strucure of the disc; regions less than 40 degrees are not shown in 3D, demonstrating the narrow central region containing hot material.}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/rotational_velocities.png}
 \label{velocity}
 \caption{Azimuthaly avereaged rotational velocity in the disc midplane with best fit curve}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/Lime_grid3.png}
 \label{grid}
 \caption{A plot of the points selected by the gridding process and the paths down which photons can propagate overlaid on a smoothed density model. The points are more concentrated at small radii and in the densest regions.}
\end{figure}

@@

Figures 6,7 and 8 show synthetic images of the CO J=3-2 line at 345.8$\,$GHz. As the upper level for this line is at 33$\,$K above the ground state this transition should be excited through out disc. given the large and uniform nature of the abundance of CO in the disc this line will be optically thick and this is only mapping the outer regions of the disc. In common with most transitions simulated, the line is seen almost exclusively in absorption of the disc continuum. The region of emission in the centre is where there is no continuum background to be absorbed.  Some indications of spiral structure can be seen but are indistinct. ***not sure how usefull these are, might be worth cutting them?***\newline

@@

The OCS lines in alma band 7 (22-21 through 30-29) with upper energy levels between 161 and 271 K above the ground state provide a way to trace hot shocked gas in spiral arms without resolving structure. Figures \ref{OCS_mom0} \& \ref{smoothOCS_mom0} show the integrated intensity of the OCS J=28-27 line in the model described previously and one where the disc section of the model is an axis-symmetric model created by taking the azimuthal average of density, temperature and molecular abundance. The order of magnitude difference in the line intensity provides a method to detect hot dense gas without being able to reolve the spiral structure spatially.\newline

@@

(figure of temperature as a function of line ratio for 2 sets of formaldehyde lines and map of derived temperature, talk to P about this. This doesnt seem doable with the lines around 1mm for alma)\newline

@@

 \caption{HCO$^+$ 3-2 PV through centre with rotation curve from figure \ref{velocity}}
\end{figure}

@@

In all the simulations except HCO$^+$ we see molecular lines in absorption throughout the majority of the disc and in some cases a small amount of emission towards the centre, where the inclination of the disc allows us the view the low density but moderate temparture region inside the inner edge of the spiral structures.\newline

@@

In this paper we have performed radiative transfer simulations of a hybrid model comprising a 0.4$\, M_\odot$ self gravitating disc with radius 64$\,AU$ showing spiral density waves, surrounded by an envelope simulated as a collapsing 10$\,M_\odot$ BE-sphere. CASA simulations show that at a distance of 100$\,pc$ both spatial resolution of the spirals in such a disc and extraction of kinematic and temperature from molecular lines are possible in ALMA band 7. Our simulations show that many  molecular spieces are prodominantly seen in absorbtion towards the centre of  self-gravitating protoplanetary discs.\newline
The quiesient nature of the envelope around such discs only obscures them within $\pm\,0.5 km\,s^{-1}$.\newline
Hot, shocked gas can be detected in such a disc without resolving it by observing OCS transitions between 26-27 to 30-29.\newline
Mention that dust and gas are assumed to be in temperature equilibrium if they arent and dust is significantly cooler then transitions may not show up in absorption. atm spirals shock heat the mid-plane layer and the colder more difusse gas absorbs the continuum from it (need to ask aaron whether this is the case, the narrow midplane is much hotter than the rest but should check that this is due to shock heating, though with there being no accreation or iradiation i dont see what else it could be)\newline
note that the disc mass is on the high end for early discs\newline
rotation curves can be gatherered from these observations even with spiral structure\newline

>>>>

%-describe the envelope structure (pre-stellar core)
The model used to simulate the emission from a young protoplanetary disc is a hybrid model created from hydrodynamical models of an infalling prestellar core and a gravitationally unstable protoplanetary disc. The prestellar core model is based upon the model of Keto \& Caselli (2010). This describes an infalling pre-stellar core, with densities temperatures and velocities given by the model of the collapse of a 10 solar mass Bonnor-Ebert sphere and provides a match to line profiles seen in the pre-stellar core L1544. Some slight modification due to the addition of Oxygen lines (Keto et al 2012 in prep) have been been added to this model resulting in a slight decrease in the peak infall velocity. This prestellar core model extends from 80$\,$au out to 10,000$\,$AU. The model is a 1D spherically symmetric model with inward motions (see figure \ref{l1544model}).\newline

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/rhoT2.png}
 \label{rhoT} 
 \caption{Left: A 3D plot of log density (m$^3$) showing the spiral structure in the xy plane and scale height of the disc. Right: The 3D temperature structure of the disc; regions less than 40 degrees are not shown in 3D, demonstrating the narrow central region containing hot material.}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/rotational_velocities.png}
 \label{velocity}
 \caption{Azimuthaly averaged rotational velocity in the disc mid-plane with best fit curve}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/Lime_grid3.png}
 \label{grid}
 \caption{A plot of the points selected by the griding process and the paths down which photons can propagate overlaid on a smoothed density model. The points are more concentrated at small radii and in the densest regions.}
\end{figure}

@@

Figures 6,7 and 8 show synthetic images of the CO J=3-2 line at 345.8$\,$GHz. As the upper level for this line is at 33$\,$K above the ground state this transition should be excited through out disc. given the large and uniform nature of the abundance of CO in the disc this line will be optically thick and this is only mapping the outer regions of the disc. In common with most transitions simulated, the line is seen almost exclusively in absorption of the disc continuum. The region of emission in the centre is where there is no continuum background to be absorbed.  Some indications of spiral structure can be seen but are indistinct. ***not sure how useful these are, might be worth cutting them?***\newline

@@

The OCS lines in alma band 7 (22-21 through 30-29) with upper energy levels between 161 and 271 K above the ground state provide a way to trace hot shocked gas in spiral arms without resolving structure. Figures \ref{OCS_mom0} \& \ref{smoothOCS_mom0} show the integrated intensity of the OCS J=28-27 line in the model described previously and one where the disc section of the model is an axis-symmetric model created by taking the azimuthal average of density, temperature and molecular abundance. The order of magnitude difference in the line intensity provides a method to detect hot dense gas without being able to resolve the spiral structure spatially.\newline

@@

(figure of temperature as a function of line ratio for 2 sets of formaldehyde lines and map of derived temperature, talk to P about this. This doesn't seem doable with the lines around 1mm for alma)\newline

@@

 \caption{HCO$^+$ 3-2 PV through centre with rotation curve from figure \ref{velocity} *** should the velocities in this be reduced by cos(30) as we're not in the plane of rotation? ***}
\end{figure}

@@

In all the simulations except HCO$^+$ we see molecular lines in absorption throughout the majority of the disc and in some cases a small amount of emission towards the centre, where the inclination of the disc allows us the view the low density but moderate temperature region inside the inner edge of the spiral structures.\newline

@@

In this paper we have performed radiative transfer simulations of a hybrid model comprising a 0.4$\, M_\odot$ self gravitating disc with radius 64$\,AU$ showing spiral density waves, surrounded by an envelope simulated as a collapsing 10$\,M_\odot$ BE-sphere. CASA simulations show that at a distance of 100$\,pc$ both spatial resolution of the spirals in such a disc and extraction of kinematic and temperature from molecular lines are possible in ALMA band 7. Our simulations show that many  molecular species are predominantly seen in absorption towards the centre of  self-gravitating protoplanetary discs.\newline
The quiescent nature of the envelope around such discs only obscures them within $\pm\,0.5 km\,s^{-1}$.\newline
Hot, shocked gas can be detected in such a disc without resolving it by observing OCS transitions between 26-27 to 30-29.\newline
Mention that dust and gas are assumed to be in temperature equilibrium if they aren't and dust is significantly cooler then transitions may not show up in absorption. atm spirals shock heat the mid-plane layer and the colder more diffuse gas absorbs the continuum from it (need to ask Aaron whether this is the case, the narrow mid-plane is much hotter than the rest but should check that this is due to shock heating, though with there being no accretion or irradiation i don't see what else it could be)\newline
note that the disc mass is on the high end for early discs\newline
rotation curves can be gathered from these observations even with spiral structure\newline
